### 2017-01-27 manuscript.rmd
updated md after merge
added final paragraph to the conclusion

Most CFG studies to date compare gene expression across species, often between distinct cells, tissues or developmental time points. Gene expression has been a popular target of CFG studies because, with the advent of RNA-Seq and related transcriptome sequencing methods, it is the most readily collected functional genomics trait across diverse species. Here we reconsider two recent comparative analyses of gene expression data [@KryuchkovaMostacci:2016iw; @Levin:2016bp].

@@

>>>>

Most CFG studies to date compare gene expression across species, often between distinct tissues or developmental time points. Gene expression has been a popular target of CFG studies because, with the advent of RNA-Seq and related transcriptome sequencing methods, it is the most readily collected functional genomics trait across diverse species. Here we reconsider two recent comparative analyses of gene expression data [@KryuchkovaMostacci:2016iw; @Levin:2016bp].

@@

These analyses also illustrate the importance of not conflating evolutionary processes with the patterns they generate. Stating that there is a pattern wherein paralogs tend to be more different than orthologs does not mean that there is a difference in the process by which orthologs and paralogs evolve. When paralogs tend to be more distantly related than orthologs, they can be more different from wach other when both evolve under the same processes. Noting that there are multiple pairwise comparisons between distantly related animals that show lower similarity in comparison at the midpoint of development does not mean that there is a general process by which this phase of development evolved deep in the animal tree. Changes restricted to a single branch will be included in multiple pairwise comparisons, and can give the false impression of a general pattern. In both cases, there is no way to interpret the impications of patterns between living organisms for understanding evolutionary processes that happened in the past without taking the relationships between the organisms - the phylogenetic tree - into account.
